---
title: "Big Data Econometrics"
subtitle: "2nd homework assignment"
author: "mansok (2017042)"
date: "`r Sys.Date()`"
output:
  rmdformats::downcute:
    highlight: tango
    code_folding: hide
    lightbox: true
    gallery: true
---

```{r setup, include=FALSE}
# I10.9, I4.13, I9.5
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=TRUE,
               cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)

# libraries
library(data.table)
library(kableExtra)
library(pander)
library(plotly)

library(ISLR)

library(MASS)
library(class)

library(e1071)

# for bonus tasks
library(tm)
library(pdftools)
library(wordcloud)

library(hdm)

panderOptions('knitr.auto.asis', FALSE)

# load comments
source("comments.R")
```



# I10.9
`r i10.9.description`

```{r dataprep1}
set.seed(1995)
dat1 <- USArrests
```

## (a)
```{r i10.9.a, results='asis'}
fit.hclust <- hclust(dist(dat1), method = "complete")
plot(fit.hclust, main = "States clusters")
```

`r i10.9.a`

## (b)
```{r i10.9.b_1, results='asis'}
fit.hclust.c <- cutree(fit.hclust, 3)
fit.hclust.c

fit.hclust.c %>% 
  table() %>% 
  t() %>% 
  # xtable() %>%
  kable(caption = "Number of states in each cluster") %>%
  kable_styling(position = "left", bootstrap_options = "hover")
```

`r i10.9.b`

## (c)
```{r i10.9.c, results='asis'}
dat2 <- scale(dat1)
fit.hclust.s <- hclust(dist(dat2), method = "complete")
plot(fit.hclust.s)
```

`r i10.9.c`

## (d)
```{r i10.9.d_1, results='asis'}
fit.hclust.s.c <- cutree(fit.hclust.s, 3)
fit.hclust.s.c

fit.hclust.s.c %>% 
  table() %>% 
  t() %>% 
  # xtable() %>%
  kable(caption = "Number of states in each cluster (scaled data)") %>%
  kable_styling(position = "left", bootstrap_options = "hover")
```

```{r i10.9.d_3, results='asis'}
dd <- table(fit.hclust.c, fit.hclust.s.c)
rownames(dd) <- c("Non-scaled 1", "Non-scaled 2", "Non-scaled 3")
colnames(dd) <- c("Scaled 1", "Scaled 2", "Scaled 3")
dd %>% 
  # xtable() %>%
  kable(caption = "Model comparison on non-scaled vs scaled data") %>%
  kable_styling(position = "left", bootstrap_options = "hover") %>% 
  column_spec(1, bold = TRUE)
```

`r i10.9.d`



# I4.13

`r i4.13.description`

## General information

```{r i4.13_1, results='asis'}
dat <- Boston
setDT(dat)
dat[ , crim.ind := fifelse(crim < median(crim), 0, 1)][ , crim := NULL]
sapply(dat, summary) %>% 
  t() %>% 
  kable(caption = "General summary of Boston dataset variables") %>% 
  kable_styling(position = "left", bootstrap_options = c("striped", "hover")) %>% 
  column_spec(1, bold = T)
```

```{r i4.13_2, results='asis'}
dat[ , .N, crim.ind] %>% 
  kable(caption = "Crime rate comparison to median") %>%
  kable_styling(position = "left", bootstrap_options = "hover")
```

## Create train and test sets
```{r i4.13_3, results='asis'}
set.seed(1995)
train_ind <- sample(nrow(dat), nrow(dat)*0.7)
train <- as.data.table(dat[train_ind, ])
test <- as.data.table(dat[-train_ind, ])

train[ , .N, crim.ind] %>% 
  kable(caption = "Crime rate comparison to median (train set)") %>%
  kable_styling(position = "left", bootstrap_options = "hover")

test[ , .N, crim.ind] %>% 
  kable(caption = "Crime rate comparison to median (test set)") %>%
  kable_styling(position = "left", bootstrap_options = "hover")
```

## Logistic regression {.tabset}

### All variables
```{r i4.13_4, results='asis'}
fit.glm <- glm(crim.ind ~ ., data = train, family = binomial)
test$predict <- predict(fit.glm, test, type = "response")
test[ , crim.p := fifelse(predict < 0.5, 0, 1)]

table(test$crim.ind, test$crim.p) %>% 
  # xtable() %>%
  kable(caption = "Actual vs predicted") %>%
  kable_styling(position = "left", bootstrap_options = "hover") %>% 
  column_spec(1, bold = T)

acc <- round(mean(test$crim.p == test$crim.ind), 3) * 100
err <- round(mean(test$crim.p != test$crim.ind), 3) * 100
cat("Test accuracy:", acc, "%; \n",
    "Test error rate:", err, "% \n")

fit.glm %>% 
  summary() %>%
  pander()
```

### Excluding non significant variables
```{r i4.13_5, results='asis'}
fit.glm2 <- glm(crim.ind ~ . - indus - chas - rm - lstat, data = train, family = binomial)
test$predict <- predict(fit.glm2, test, type = "response")
test[ , crim.p := fifelse(predict < 0.5, 0, 1)]

table(test$crim.ind, test$crim.p) %>% 
  # xtable() %>%
  kable(caption = "Actual vs predicted") %>%
  kable_styling(position = "left", bootstrap_options = "hover") %>% 
  column_spec(1, bold = T)

acc <- round(mean(test$crim.p == test$crim.ind), 3) * 100
err <- round(mean(test$crim.p != test$crim.ind), 3) * 100
cat("Test accuracy:", acc, "%; \n",
    "Test error rate:", err, "% \n")

fit.glm2 %>% 
  summary() %>%
  pander()
```


## LDA {.tabset}

### All variables
```{r i4.13_6, results='asis'}
fit.lda <- lda(crim.ind ~ ., data = train)
test$crim.p <- predict(fit.lda, test, type = "response")$class

table(test$crim.ind, test$crim.p) %>% 
  # xtable() %>%
  kable(caption = "Actual vs predicted") %>%
  kable_styling(position = "left", bootstrap_options = "hover") %>% 
  column_spec(1, bold = T)

acc <- round(mean(test$crim.p == test$crim.ind), 3) * 100
err <- round(mean(test$crim.p != test$crim.ind), 3) * 100
cat("Test accuracy:", acc, "%; \n",
    "Test error rate:", err, "% \n")

fit.lda %>% 
  summary() %>%
  pander()
```

### Excluding some variables
```{r i4.13_7, results='asis'}
fit.lda2 <- lda(crim.ind ~ . - indus - chas - rm - lstat, data = train)
test$crim.p <- predict(fit.lda2, test, type = "response")$class

table(test$crim.ind, test$crim.p) %>% 
  # xtable() %>%
  kable(caption = "Actual vs predicted") %>%
  kable_styling(position = "left", bootstrap_options = "hover") %>% 
  column_spec(1, bold = T)

acc <- round(mean(test$crim.p == test$crim.ind), 3) * 100
err <- round(mean(test$crim.p != test$crim.ind), 3) * 100
cat("Test accuracy:", acc, "%; \n",
    "Test error rate:", err, "% \n")

fit.lda2 %>% 
  summary() %>%
  pander()
```

## KNN {.tabset}
```{r i4.13_8, results='asis'}
# select features
nm <- which(!names(dat) %in%c("indus", "chas", "rm", "lstat"))
dat.knn <- copy(dat[, nm, with = FALSE])

# transform variables
nm <- names(dat.knn)
nm <- nm[!grepl("crim.ind", nm)]
dat.knn[ , (nm) := lapply(.SD, scale), .SDcols = nm]

# split train - test sets
train.knn <- dat.knn[train_ind, ]
test.knn <- dat.knn[-train_ind, ]

```

### K=1
```{r i4.13_9, results='asis'}
# KNN with K=1
crim.p <- knn(train.knn[ , !"crim.ind"], test.knn[ , !"crim.ind"], train.knn[ , crim.ind], k = 1)
table(test.knn[ , crim.ind], crim.p) %>% 
  # xtable() %>%
  kable(caption = "Actual vs predicted") %>%
  kable_styling(position = "left", bootstrap_options = "hover") %>% 
  column_spec(1, bold = T)

acc <- round(mean(crim.p == test.knn$crim.ind), 3) * 100
err <- round(mean(crim.p != test.knn$crim.ind), 3) * 100
cat("Test accuracy:", acc, "%; \n",
    "Test error rate:", err, "% \n")
```

### K=3
```{r i4.13_10, results='asis'}
# KNN with K=3
crim.p3 <- knn(train.knn[ , !"crim.ind"], test.knn[ , !"crim.ind"], train.knn[ , crim.ind], k = 3)
table(test.knn[ , crim.ind], crim.p3) %>% 
  # xtable() %>%
  kable(caption = "Actual vs predicted") %>%
  kable_styling(position = "left", bootstrap_options = "hover") %>% 
  column_spec(1, bold = T)

acc <- round(mean(crim.p3 == test.knn$crim.ind), 3) * 100
err <- round(mean(crim.p3 != test.knn$crim.ind), 3) * 100
cat("Test accuracy:", acc, "%; \n",
    "Test error rate:", err, "% \n")
```

### K=5
```{r i4.13_11, results='asis'}
# KNN with K=5
crim.p5 <- knn(train.knn[ , !"crim.ind"], test.knn[ , !"crim.ind"], train.knn[ , crim.ind], k = 5)
table(test.knn[ , crim.ind], crim.p5) %>% 
  # xtable() %>%
  kable(caption = "Actual vs predicted") %>%
  kable_styling(position = "left", bootstrap_options = "hover") %>% 
  column_spec(1, bold = T)

acc <- round(mean(crim.p5 == test.knn$crim.ind), 3) * 100
err <- round(mean(crim.p5 != test.knn$crim.ind), 3) * 100
cat("Test accuracy:", acc, "%; \n",
    "Test error rate:", err, "% \n")
```

`r i4.13.final`

# I9.5

`r i9.5.description`

## (a)
```{r i9.5.a, results='asis'}
set.seed(1995)
x1 <- runif(500) - 0.5
x2 <- runif(500) - 0.5
y <- 1 * (x1^2 - x2^2 > 0)
```

## (b)
```{r i9.5.b, results='asis'}
dat9 <- data.table(as.factor(y), x1, x2)
# plot_ly(dd, x=~x1) %>% add_markers(y = ~x2, showlegend = FALSE)


fig <- plot_ly(dat9, x = ~x1, y = ~x2, color = ~y, symbol = ~y,
               type = 'scatter', mode = 'markers', hoverinfo = 'text',
               marker = list(size = 7),
               text = ~paste('</br> Y class: ', y,
                             '</br> X1: ', round(x1, 3),
                             '</br> X2: ', round(x2, 3)))

fig <- fig %>% layout(title = 'X1 vs X2',
                      xaxis = list(title = 'X1'),
                      yaxis = list(title = 'X2'))
fig
```

## (c)
```{r i9.5.c, results='asis'}
fit.glm9 <- glm(y ~ x1 + x2, family = binomial)
fit.glm9 %>% 
  summary()  %>%
  pander()
```

## (d)
```{r i9.5.d, results='asis'}
dat9$predict <- predict(fit.glm9, dat9, type = "response")
dat9[ , y.p := fifelse(predict < 0.5, 0, 1)]

fig <- plot_ly(dat9, x = ~x1, y = ~x2, color = ~y.p, symbol = ~y.p,
               type = 'scatter', mode = 'markers', hoverinfo = 'text',
               marker = list(size = 7),
               text = ~paste('</br> Real: ', y,
                             '</br> Predicted: ', y.p,
                             '</br> X1: ', round(x1, 3),
                             '</br> X2: ', round(x2, 3)))

fig <- fig %>% layout(title = 'Logistic regression results',
                      xaxis = list(title = 'X1'),
                      yaxis = list(title = 'X2'))
fig
```

## (e)
```{r i9.5.e, results='asis'}
fit.glm92 <- glm(y ~ x1 + x2 + I(x1^2) + I(x2^2) + I(x1 * x2), data = dat9, family = binomial)
fit.glm92 %>% 
  summary()  %>%
  pander()
```

## (f)
```{r i9.5.f, results='asis'}
dat9$predict2 <- predict(fit.glm92, dat9, type = "response")
dat9[ , y.p2 := fifelse(predict2 < 0.5, 0, 1)]

fig <- plot_ly(dat9, x = ~x1, y = ~x2, color = ~y.p2, symbol = ~y.p2,
               type = 'scatter', mode = 'markers', hoverinfo = 'text',
               marker = list(size = 7),
               text = ~paste('</br> Real: ', y,
                             '</br> Predicted: ', y.p2,
                             '</br> X1: ', round(x1, 3),
                             '</br> X2: ', round(x2, 3)))

fig <- fig %>% layout(title = 'Logistic regression results',
                      xaxis = list(title = 'X1'),
                      yaxis = list(title = 'X2'))
fig
```

## (g)
```{r i9.5.g, results='asis'}
fit.svm <- svm(as.factor(y) ~ x1 + x2, dat9, kernel = "linear", cost = 0.01)
dat9$y.p3 <- predict(fit.svm, dat9, type = "response")

fig <- plot_ly(dat9, x = ~x1, y = ~x2, color = ~y.p3, symbol = ~y.p3,
               type = 'scatter', mode = 'markers', hoverinfo = 'text',
               marker = list(size = 7),
               text = ~paste('</br> Real: ', y,
                             '</br> Predicted: ', y.p3,
                             '</br> X1: ', round(x1, 3),
                             '</br> X2: ', round(x2, 3)))

fig <- fig %>% layout(title = 'SVM results',
                      xaxis = list(title = 'X1'),
                      yaxis = list(title = 'X2'))
fig
```

## (h)
```{r i9.5.h, results='asis'}
fit.svm2 = svm(as.factor(y) ~ x1 + x2, dat9, gamma = 1)
dat9$y.p4 <- predict(fit.svm2, dat9, type = "response")

fig <- plot_ly(dat9, x = ~x1, y = ~x2, color = ~y.p4, symbol = ~y.p4,
               type = 'scatter', mode = 'markers', hoverinfo = 'text',
               marker = list(size = 7),
               text = ~paste('</br> Real: ', y,
                             '</br> Predicted: ', y.p4,
                             '</br> X1: ', round(x1, 3),
                             '</br> X2: ', round(x2, 3)))

fig <- fig %>% layout(title = 'SVM results',
                      xaxis = list(title = 'X1'),
                      yaxis = list(title = 'X2'))
fig
```


# Bonus Tasks

## 1. Text document analysis {.tabset}

`r bonus1.description`

### Words frequency
```{r bonus1.1, results='asis'}
dat.pdf <- Corpus(URISource("straipsnis.pdf"),
                  readerControl = list(reader = readPDF))

dtm <- TermDocumentMatrix(dat.pdf,
                          control = 
                            list(removePunctuation = TRUE,
                                 stopwords = TRUE,
                                 tolower = TRUE,
                                 stemming = TRUE,
                                 removeNumbers = TRUE,
                                 bounds = list(global = c(1, Inf)))) 

mm <- as.matrix(dtm)
ww <- sort(rowSums(mm), decreasing = TRUE)
dd <- data.frame(word = names(ww), freq = ww)

head(dd, 10) %>% 
  kable(caption = "10 most popular words") %>%
  kable_styling(position = "left", bootstrap_options = "hover")
```

### Words frequency plot
```{r bonus1.2, results='asis'}
set.seed(1995)
wordcloud(words = dd$word, freq = dd$freq, min.freq = 5,
          max.words = 200, random.order = FALSE, rot.per = 0.35, 
          colors = brewer.pal(8, "Dark2")
)
```

`r bonus1.paper`

### Words frequency plot by section
```{r bonus1.3, results='asis'}
dat.pdf.c <- VCorpus(VectorSource(pdf_text("straipsnis.pdf")))

dtm.c <- TermDocumentMatrix(dat.pdf.c, 
                            control = 
                              list(removePunctuation = TRUE,
                                   stopwomm2rds = TRUE,
                                   tolower = TRUE,
                                   stemming = TRUE,
                                   removeNumbers = TRUE,
                                   bounds = list(global = c(1, Inf))))

mm2 <- as.matrix(dtm.c)

for(i in 1:(ncol(mm2))){
  stat <- mm2[,i]
  v <- sort(stat,decreasing=TRUE)
  d <- data.frame(word = names(v), freq = v)
  
  wordcloud(words = d$word, freq = d$freq, min.freq = 1,
            max.words = 20, random.order = FALSE, rot.per = 0.35, 
            colors = brewer.pal(8, "Dark2"))
}
```

`r bonus1.sub`

## 2. Barro-Lee example {.tabset}

`r bonus2.description`

### Dropping 20 countries
```{r bonus2.1, results='asis'}
set.seed(19950330)

dat.b <- GrowthData
setDT(dat.b)
drop.ind <- sample(1:(nrow(dat.b)), 20)
dat.b <- dat.b[!drop.ind, ]

dim.b <- dim(dat.b)
cat("Data dimensions after dropping 20 countries: rows:", dim.b[1], "; columns:", dim.b[2], ". \n")
```

```{r bonus2.2, results='asis'}
# from vignette
y <- dat.b[, 1, drop = F]
d <- dat.b[, 3, drop = F]
X <- as.matrix(dat.b)[, -c(1, 2, 3)]
varnames <- colnames(dat.b)

# OLS:
xnames <- varnames[-c(1, 2, 3)] 
dandxnames <- varnames[-c(1, 2)] 

# create formulas by pasting names (this saves typing times)
fmla <- as.formula(paste("Outcome ~ ", paste(dandxnames, collapse = "+")))
ls.effect <- lm(fmla, data = dat.b)

# effect by the partialling out by Post-Lasso:
lasso.effect <- rlassoEffect(x = X, y = y, d = d, method = "partialling out")
summary(lasso.effect)
```

```{r bonus2.3, results='asis'}
# double selection:
doublesel.effect <- rlassoEffect(x = X, y = y, d = d, method = "double selection")
summary(doublesel.effect)
```

```{r bonus2.4, results='asis'}
table <- rbind(summary(ls.effect)$coef["gdpsh465", 1:2], 
               summary(lasso.effect)$coef[, 1:2],
               summary(doublesel.effect)$coef[, 1:2])
colnames(table) <- c("Estimate", "Std. Error")
rownames(table) <- c("OLS",
                     "Post-Lasso ",
                     "Double selection")

table %>% 
  # xtable(digits=c(2, 2,5))%>% 
  # xtable() %>%
  kable(caption = "Models comparison") %>%
  kable_styling(position = "left", bootstrap_options = "hover") %>% 
  column_spec(1, bold = T)
```


### Full dataset
```{r bonus2.5, results='asis'}
dat.b.f <- GrowthData
setDT(dat.b.f)

dim.b.f <- dim(dat.b.f)
cat("Full dataset dimensions: rows:", dim.b.f[1], "; columns:", dim.b.f[2], ". \n")
```

```{r bonus2.6, results='asis'}
# from vignette
y <- dat.b.f[, 1, drop = F]
d <- dat.b.f[, 3, drop = F]
X <- as.matrix(dat.b.f)[, -c(1, 2, 3)]
varnames <- colnames(dat.b.f)

# OLS:
xnames <- varnames[-c(1, 2, 3)] 
dandxnames <- varnames[-c(1, 2)] 

# create formulas by pasting names (this saves typing times)
fmla <- as.formula(paste("Outcome ~ ", paste(dandxnames, collapse = "+")))
ls.effect <- lm(fmla, data = dat.b.f)

# effect by the partialling out by Post-Lasso:
lasso.effect <- rlassoEffect(x = X, y = y, d = d, method = "partialling out")
summary(lasso.effect)
```

```{r bonus2.7, results='asis'}
# double selection:
doublesel.effect <- rlassoEffect(x = X, y = y, d = d, method = "double selection")
summary(doublesel.effect)
```

```{r bonus2.8, results='asis'}
table <- rbind(summary(ls.effect)$coef["gdpsh465", 1:2], 
               summary(lasso.effect)$coef[, 1:2],
               summary(doublesel.effect)$coef[, 1:2])
colnames(table) <- c("Estimate", "Std. Error")
rownames(table) <- c("OLS",
                     "Post-Lasso ",
                     "Double selection")

table %>% 
  kable(caption = "Models comparison") %>%
  kable_styling(position = "left", bootstrap_options = "hover") %>% 
  column_spec(1, bold = T)
```